---
title: "Introduction to R and Shiny"
author: "Renato A. Correa dos Santos"
date: "2022-12-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R, Markdown, and Shiny

### R

#### Why R?

R was released in the early 1990s by Ross Ihaka and Robert Gentleman, in Australia, as part of the Open Source movement.

* Powerful for data analysis and graphical tools
* Free
* Availability of libraries

#### Variables



#### Basic R data types

##### Characters (non-numeric values)

Characters represent textual information, or strings (this is how we can textual information in Python, for example).

Characters could be used as categorical variables; however, R (as a language strongly used in statistics) has the `factor` data structure that is more suitable for this purpose.

```{r}
# Creating a string
sentence  <- "I love Systems Biology"
sentence
```

Let's see what we could do with characters.

```{r}

```

##### Numeric (real or decimal)



##### Integer



##### Logical (non-numeric values)



##### Complex




#### Basic R data structures

##### Vectors



##### Matrices



##### Lists



##### Data frames



##### Factors

Depending on the reference, it can be considered as an R data type (e.g. The Book of R by TM Davies)


#### Reading and writing files

I science, we usually have tabulated text files or Excel spreasheets that we use for data analysis. We can import these files into R as well. In our course, we are also going to import files that store network information.

To read files, we can use:

* `read.csv` to read a file with comma-separated values
* `read.table` to read a file with tabulation-separeted values

As mentioned in [here](http://swcarpentry.github.io/r-novice-inflammation/11-supp-read-write-csv/), there are libraries specific for handling Excel spreadsheets. However, we are going to use only CSV and TSV files in this course. 

Eventually, we may want to write data into a file.

* `write.csv`

#### Plotting




### RMarkdown



### Shiny


## Bibliography

* [Programming with R (Software Carpentry Lesson)](http://swcarpentry.github.io/r-novice-inflammation/)