---
title: "Introduction to network models"
author: "Renato A. Correa dos Santos"
date: "2022-11-29"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to network models

Theory (slides):

* Reductionism
* Holism
* Complex systems
* Brief history of graphs
  * e.g., Euler and the Königsberg’s bridge paradigm
  * e.g., Kochen and Pool and the structure of human society
  * e.g., Erdos and Rényi and the "Small World" effect
  * e.g., Barabási and Albert and the "Scale-free" graphs
* Network definition
* Classification of networks
  * Nature of ties: directed and undirected
  * Ties can be binary (unweighted) and have an associated value (weighted)
  * Cyclic and acyclic ?
  * Types of nodes: Unipartite (monopartite or one-mode), bipartite (two-mode)
  * Multi-edge
  * Tree: acyclic undirected graph
  * Connected graph: there is a path from any point to any other point
  * Complete graph: every pair of distinct vertices is connected by a unique edge
  * Cluster: graph formed from the disjoint union of complete graphs
  * Clique: subset of vertices such that every pair is connected
* Network properties and metrics
  * Degree: total number of edges adjancent to a vertex
  * Density: ratio between number of edge and the possible number of edges in the same graph (density allows to classify graphs in dense and sparse)
  * Clustering coefficient: 
  * Matching index: 
  * Distance: length of the shortest path between given nodes
  * Network centrality: 
    * Closeness centrality:
    * Betweenness centrality: 
    * Eccentricity centrality:
  * Motifs: repeated graphlets (small subgraphs of a larger network)
    * 
* R tools for network analysis
  * [statnet suite](https://statnet.org/) and its [packages](https://statnet.org/packages/)
  * [igraph package](https://github.com/igraph/igraph)
  * [tidygraph package](https://github.com/thomasp85/tidygraph)
  * [ggraph package](https://github.com/thomasp85/ggraph)

## Practical session on network models

### Data structures



#### Creating a first network in R

First test ...

```{r}
#install.packages("tidyverse")
library(tidyverse)

# Molecules in Fig.3A of Amores and Martínez-Antonio (2022)
edge_list <- tibble(from = c("molecule1", "molecule1", "molecule2", "molecule2"), to = c("molecule2", "molecule4", "molecule3", "molecule4"))
edge_list
```

Creating an undirected network:

```{r}
#install.packages("network")
library(network)
molecules_network <- network(edge_list,
                             directed = FALSE)
molecules_network
```

#### Edge list



### File representation of networks

There are several ways we can store network information in files.

* Adjacency matrix
* Edge list
* Node list

#### Importing files


### Network properties

#### Network centralities

* Degree centrality

* Closeness centrality
* Betweenness centrality 
* Eccentricity centrality







### Modelling




### Overview of all functions used in this practical session

* Tidyverse
* network
* 

## Bibliography

* [Interesting post "Network Analysis with R" by Jesse Sadler](https://www.jessesadler.com/post/network-analysis-with-r/)
* [Amores and Martínez-Antonio, 2022](https://doi.org/10.1007/s10142-022-00907-y)
* [Koutrouli et al. 2020](https://www.frontiersin.org/articles/10.3389/fbioe.2020.00034/full)
