---
title: "Introduction to network models"
author: "Renato A. Correa dos Santos"
date: "2022-11-29"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to network models

Theory (slides):

* Reductionism
* Holism
* Complex systems
* Brief history of graphs
  * e.g., Euler and the Königsberg’s bridge paradigm
  * e.g., Kochen and Pool and the structure of human society
  * e.g., Erdos and Rényi and the "Small World" effect
  * e.g., Barabási and Albert and the "Scale-free" graphs
* Network definition
* Classification of networks
  * Directed and undirected
  * Weighted and unweighted
  * Cyclic and acyclic ?
  * Uni-, bi-partite ?
  * 
* R tools for network analysis
  * [statnet suite](https://statnet.org/) and its [packages](https://statnet.org/packages/)
  * [igraph package](https://github.com/igraph/igraph)
  * [tidygraph package](https://github.com/thomasp85/tidygraph)
  * [ggraph package](https://github.com/thomasp85/ggraph)

## Practical session on network models

### File representation of networks

* Adjacency matrix
* Edge list
* Node list

```{r}

```



### Overview of all functions used in this practical session



## Bibliography

* [Interesting post "Network Analysis with R" by Jesse Sadler](https://www.jessesadler.com/post/network-analysis-with-r/)
